\section{Introduction}

A foundational bottleneck in large language model (LLM) inference is the sequential, left-to-right nature of autoregressive decoding. While causal masking is fundamental to this paradigm, it forces a trade-off between latency and reasoning depth: complex tasks require long chains of thought, but long chains incur prohibitive latency.

Recent research has attempted to bypass this bottleneck through ``around-the-model'' parallelization. Techniques like Skeleton-of-Thought (SoT) \cite{ning2023skeleton} decompose tasks into sub-questions and prompt the model to answer them in parallel. While promising, these methods treat the model as a black box, leading to a critical failure mode we term \textbf{Coherence Drift}: because parallel streams cannot communicate, they often hallucinate conflicting facts or redundant content.

We address this by moving both planning and coordination \textit{inside} the model. We introduce the \textbf{Parallel Decoder Transformer (PDT)}, a decoder-only architecture whose central mechanism is a mandatory prompt-time latent planning pass followed by a shared embeddings-only coordination bus. The planner seeds a shared latent state at $t=0$, and downstream streams condition on that state through SNC rather than waiting for a textual plan string to be written into the prompt or passed between streams. The resulting coordination policy is implemented by a stack rather than a single head: the planner provides the initial latent code, while plan projection, note emission, SNC consumption, and rollback heads maintain the shared state during decoding.

\subsection{The Efficiency Imperative}
Prior attempts at architectural modification often require expensive full-model fine-tuning, which is computationally prohibitive for modern 70B+ parameter models. A key design constraint of PDT is therefore that the base language model's weights remain frozen while coordination is added through sidecar modules.

We present a \textbf{Parameter-Efficient} design where the massive transformer ``trunk'' remains frozen and coordination is delegated to lightweight \textit{Stream Adapters}, planner modules, and SNC heads. This approach allows PDT to be deployed as a ``sidecar'' to existing open-weights models without changing the core causal decoder.

\subsection{Our Contributions}
Our contributions are as follows:
\begin{enumerate}
    \item \textbf{Planner-first latent coordination stack.} PDT performs a prompt-time planner pass that predicts fixed latent plan slots over a shared plan vocabulary, re-embeds those slots into notes space, and broadcasts the resulting seed as snapshot 0 before any stream emits tokens; later modules then consume, refresh, and police that workspace during decoding.
    \item \textbf{Versioned embedding bus semantics.} SNC reads lagged note embeddings every decode step, while cadence and agreement logic control when new snapshots are written or when rollback is triggered. Textual plans and notes remain auxiliary views for data construction, scoring, and debugging rather than the representation carried across streams.
    \item \textbf{Implementation-aligned dataset and training contract.} The dataset writes a real snapshot-0 plan seed, hashes canonical plan items into the planner vocabulary, and can mask LM supervision so each stream trains only on its own section.
\end{enumerate}

Because the coordination stack has been materially revised to match this planner-first latent contract, previously reported quantitative results are omitted from this version of the paper. Section \ref{sec:experiments} therefore focuses on the evaluation scope and the metrics that matter for the revised system, including the distinction between stack-level mechanism claims and head-level attribution claims.

The rest of the paper is organized as follows. Section \ref{sec:related} reviews related work in parallel decoding, shared workspaces, and latent coordination. Section \ref{sec:architecture} details the frozen-trunk architecture and planner-first curriculum. Section \ref{sec:snc} gives an implementation-grounded formalism for the latent planner and Dynamic Notes Bus runtime. Section \ref{sec:experiments} defines the current evaluation scope for the revised coordination stack.
