\section{Related Work}

\subsection{Token-Level Acceleration Versus Coordination}


Standard autoregressive decoding generates tokens sequentially as $p(x_t \mid x_{<t})$. Token-level acceleration methods such as Speculative Decoding \cite{leviathan2023fast}, Blockwise Parallel Decoding \cite{Stern2018}, Medusa \cite{cai2024medusa}, EAGLE \cite{EAGLE2024}, Lookahead Decoding \cite{Li2024lookahead}, and related analyses of speculative trade-offs \cite{Yan2024,survey2025} improve local efficiency by predicting or verifying multiple future tokens. More recent work also studies concurrent attention and interleaved planning mechanisms for faster generation \cite{Rodionov2025,Xiao2025sprint}.

These methods are important reference points, but they solve a different problem. They accelerate a single causal stream or speculate over future tokens within one stream. PDT is not primarily a faster decoding scheme; it is a model-internal synchronization mechanism for deciding when multiple partial generations may safely continue in parallel.

\subsection{External Prompt-Level Decomposition}


Prompt-level approaches such as Skeleton-of-Thought (SoT) \cite{ning2023skeleton} and PSLM \cite{PSLM2024} ask a model for a decomposition and then launch multiple prompted generations. Their central advantage is semantic modularity: different calls can elaborate different parts of a solution. Their central limitation is where coordination lives. Shared state remains outside the model and is mediated by prompt text, API orchestration, or post-hoc merging. As a result, these systems can decompose work but cannot provide an internal continuation rule for when sibling streams have supplied enough information for safe further generation.

PDT is closest in motivation to this family of methods, but differs in mechanism. The decomposition prior, the shared workspace, and the continuation gate all live inside one decoder rather than in an external controller.

\subsection{Shared Workspaces, Blackboard Systems, and Latent Communication}


A growing literature coordinates multiple model calls or agents through shared workspaces, graph structures, or latent communication channels. GPTSwarm \cite{Zhuge2024GPTSwarm}, Mixture-of-Agents \cite{Wang2025MoA}, blackboard-style systems \cite{Han2025Blackboard,Salemi2025Blackboard}, and topology-adaptive agent systems such as AMAS \cite{Leong2025AMAS} all point toward coordination as a systems problem rather than a single-prompt problem. Earlier work on shared neural workspaces \cite{Goyal2021} provides a conceptual precedent for a global coordination substrate, while LatentMAS \cite{Zou2025LatentMAS} demonstrates the value of latent rather than text-mediated communication.

PDT shares that intuition but places the workspace inside a single frozen decoder and binds it directly to the output interface. Its streams are not separate prompted agents. They are coordinated generation branches of one model, each with its own local cache but a common latent synchronization workspace.

\subsection{Parameter-Efficient Adaptation and Distillation}


Because modifying all parameters of a large decoder is computationally expensive, PDT follows the parameter-efficient adaptation tradition of LoRA \cite{hu2021lora} and adapters \cite{houlsby2019parameter}. Distillation and learning-with-privileged-information frameworks \cite{LopezPaz2015,Vapnik2015} further motivate training auxiliary coordination modules with teacher signals that may be unavailable or undesirable at inference time.

These techniques matter here not only for efficiency, but because they make PDT a realistic architectural extension of a frozen model. The planner, synchronization workspace, and commit-control heads can be added without rewriting the underlying language model.
