\section{Architecture}

The Parallel Decoder Transformer turns one frozen decoder into $K$ coordinated output streams. Its central design decision is that decomposition and cross-stream state are \textbf{model-internal}. The prompt is planned once, the resulting latent plan initializes a shared workspace, and each stream decodes against that workspace rather than against a separate prompt string.

\subsection{Frozen Trunk Topology}


We initialize PDT from a pre-trained decoder-only backbone parameterized by $\theta_{\text{pre}}$ and freeze all weights in $\theta_{\text{pre}}$. Trainable parameters $\phi$ are introduced as a lightweight coordination stack. The full parameter set is
\begin{equation}
\Theta = \theta_{\text{pre}} \cup \phi.
\end{equation}

The trainable subset $\phi$ contains four components:

\begin{itemize}
    \item \textbf{Stream adapters}, inserted into selected transformer blocks to provide stream-specific conditioning.
    \item \textbf{SNC backends}, implemented as cross-attention layers that read from the shared latent workspace.
    \item \textbf{Planner and notes modules}, including a fixed-slot planner head, a shared plan-embedding matrix, and a projection into the notes space used by the bus.
    \item \textbf{Auxiliary control heads}, including note-emission, coverage, agreement, and stream-classification heads.
\end{itemize}

All streams share the same frozen trunk parameters, but each stream maintains its own KV cache, adapter state, and decode position. The base language model therefore remains canonical, while coordination is expressed through sidecar modules rather than through changes to pre-trained weights.

\subsection{Prompt-Time Latent Planner}


Let $\mathbf{H}_x \in \mathbb{R}^{T \times d}$ denote the hidden states of the prompt under the frozen trunk. PDT performs a mandatory planning pass before any output tokens are generated. The planner predicts logits over $S$ latent plan slots and a shared plan vocabulary of size $V_p$:
\begin{equation}
\mathbf{\Pi} = \mathrm{PlannerHead}(\mathbf{H}_x) \in \mathbb{R}^{S \times V_p}, \qquad
z_i = \arg\max_{a} \mathbf{\Pi}_{i,a}.
\end{equation}
In the current implementation, $S = 16$ and $V_p = 65{,}536$.

The latent slot identifiers $z_{1:S}$ are not language-model tokens. They index a shared latent plan vocabulary that is also used by the plan embedding matrix, the canonical plan catalog, and coverage targets. Active slots are re-embedded and projected into notes space:
\begin{equation}
\mathbf{e}_i = E_{\text{plan}}[z_i], \qquad
\mathbf{n}^{\text{plan}}_0 =
\mathrm{Norm}\!\left(
\frac{1}{|M|}\sum_{i \in M} W_{\text{plan}} \mathbf{e}_i
\right),
\end{equation}
where $M$ denotes the active planner slots.

The resulting vector $\mathbf{n}^{\text{plan}}_0$ is broadcast as \textbf{snapshot 0} on the Dynamic Notes Bus before any stream emits tokens. The planner does not merely assign semantic subtopics. It initializes the shared coordination state against which subsequent continuation decisions are made. Snapshot 0 therefore serves both as a decomposition prior and as the first synchronization contract among streams.

In PDT, parallel generation does not begin from independent empty states. It begins from a common latent commitment structure.

\subsection{Dynamic Notes Bus as Synchronization Workspace}


The \textbf{Dynamic Notes Bus} is PDT's shared latent workspace. It is an embeddings-only, versioned store of planner and stream summaries. Textual plans and rendered notes may be used for dataset construction, observability, or supervision, but inference-time coordination uses only note embeddings.

We describe the runtime at two timescales:

\begin{itemize}
    \item $\tau$: the number of tokens in each provisional block emitted per stream between synchronization decisions,
    \item $\Delta$: the reveal delay between a note being emitted and becoming visible to sibling streams,
    \item $H$: the rollback or commit horizon over which recently generated provisional content may be withheld or regenerated.
\end{itemize}

Let $\widehat{\mathbf{n}}^{(j)}_u$ denote the provisional note emitted by stream $j$ at synchronization round $u$, and let $B$ be the workspace budget. The visible notes window for stream $k$ at round $v$ is
\begin{equation}
\mathcal{W}^{(k)}_v =
\mathrm{Window}\!\left(
\left\{\mathbf{n}^{\text{plan}}_0\right\}
\cup
\left\{\widehat{\mathbf{n}}^{(j)}_u : j \in [K],\; u \le v-\Delta\right\},
B
\right).
\end{equation}

The Dynamic Notes Bus is therefore not just an auxiliary memory. It is the synchronization workspace that determines whether provisional multi-stream generation may be committed and extended.

\subsection{Synchronized Block Emission Protocol}


PDT runs in synchronized rounds rather than as unconstrained free-running streams. At synchronization round $v$:

\begin{enumerate}
    \item The planner-seeded bus exposes the visible notes window $\mathcal{W}^{(k)}_v$ to each active stream $k$.
    \item Each stream emits a provisional block of $\tau$ tokens conditioned on its local cache and the visible workspace:
\begin{equation}
   \widehat{\mathbf{y}}^{(k)}_v =
   \left(\hat y^{(k)}_{v,1}, \ldots, \hat y^{(k)}_{v,\tau}\right).
\end{equation}
    \item At the end of the block, stream $k$ emits a provisional latent note $\widehat{\mathbf{n}}^{(k)}_v$ summarizing what it has established, which plan items it believes it owns, and which dependencies remain unresolved.
    \item Coverage and agreement heads evaluate whether the provisional block is consistent with the shared plan and whether sibling state is sufficient for further continuation.
    \item If agreement passes, the block is committed and the new note snapshot becomes visible after delay $\Delta$.
    \item If agreement fails, the system withholds commit, stalls a subset of streams, or rolls back and regenerates within horizon $H$ using fresher shared context.
\end{enumerate}
This protocol is the core mechanism of PDT. Streams are allowed to generate in parallel only up to a provisional block boundary; continuation beyond that boundary is gated by agreement over the shared latent workspace.

\subsection{Speculative Note Conditioning Cross-Attention}


SNC is the read mechanism over the visible workspace. During token emission inside round $v$, stream $k$ conditions on $\mathcal{W}^{(k)}_v$ at every decode step. Let $\mathbf{H}^{(k)}_{v,u}$ denote the hidden state for token position $u \in \{1,\ldots,\tau\}$ in round $v$. The SNC projections are
\begin{equation}
\mathbf{Q}^{(k)}_{v,u} = \mathbf{H}^{(k)}_{v,u} \mathbf{W}_Q, \qquad
\mathbf{K}^{(k)}_v = \mathcal{W}^{(k)}_v \mathbf{W}_K, \qquad
\mathbf{V}^{(k)}_v = \mathcal{W}^{(k)}_v \mathbf{W}_V,
\end{equation}
with learned parameters $\mathbf{W}_Q, \mathbf{W}_K, \mathbf{W}_V \in \phi$.

The resulting SNC context vector $\mathbf{C}^{(k)}_{v,u}$ is injected through a trust-gated residual:
\begin{equation}
\widetilde{\mathbf{H}}^{(k)}_{v,u} =
\mathbf{H}^{(k)}_{v,u}
+
\lambda_l \cdot \mathbf{C}^{(k)}_{v,u} \mathbf{W}_O,
\end{equation}
where $\lambda_l \in [0,1]$ is a learned residual gate for the instrumented layer. The gate preserves the magnitude statistics of the frozen trunk early in training and allows coordination to become stronger only when the auxiliary path is reliable.

SNC provides continuous low-bandwidth conditioning during token emission, but synchronization decisions are made only at block boundaries. PDT therefore separates two timescales: \textbf{token-time latent conditioning} and \textbf{block-time continuation control}.

\subsection{Provisional Writes, Ownership, and Commit Readiness}


At the end of each provisional block, stream $k$ writes a latent summary back to the bus:
\begin{equation}
\widehat{\mathbf{n}}^{(k)}_v =
\mathrm{SpeculationHead}\!\left(
\widetilde{\mathbf{H}}^{(k)}_{v,\tau}
\right).
\end{equation}
That note is intended to summarize the stream's newly established content, its unresolved dependencies, and its current ownership claims.

Coverage is predicted against the planner-seeded plan catalog:
\begin{equation}
\mathbf{c}^{(k)}_v =
\mathrm{CoverageHead}\!\left(
\widetilde{\mathbf{H}}^{(k)}_{v,\tau},
E_{\text{plan}}(z_{1:S}),
m_{\text{plan}}
\right),
\end{equation}
where $m_{\text{plan}}$ masks inactive or padded plan items.

Because coverage targets are stream-qualified canonical plan items, they track more than semantic relevance. They also track ownership and non-overlap. A provisional block is not ready to commit merely because it is fluent; it must also preserve ownership consistency with respect to the planner-seeded plan catalog.

\subsection{Agreement-Gated Commit and Rollback Control}


Agreement in PDT is not only a rollback trigger after a bad block. It is the permission signal for the next parallel block. For each active stream $k$ at round $v$, PDT predicts a \textbf{readiness score}
\begin{equation}
r^{(k)}_v =
\mathrm{AgreeHead}\!\left(
\widetilde{\mathbf{H}}^{(k)}_{v,\tau},
\mathcal{W}^{(k)}_v,
\mathbf{c}^{(k)}_v,
\widehat{\mathbf{n}}^{(k)}_v
\right),
\end{equation}
which estimates whether stream $k$ has enough synchronized information about sibling streams to safely commit its current block and continue.

The simplest continuation rule is a global gate over active streams:
\begin{equation}
A_v =
\mathbf{1}\!\left[
\min_{k \in \mathcal{A}_v} r^{(k)}_v > \gamma
\right],
\end{equation}
where $\mathcal{A}_v$ is the set of active streams at round $v$ and $\gamma$ is a learned or tuned threshold.

If $A_v = 1$, PDT commits the provisional blocks $\widehat{\mathbf{y}}^{(k)}_v$ and schedules the new notes $\widehat{\mathbf{n}}^{(k)}_v$ to become visible after delay $\Delta$. If $A_v = 0$, the default policy is to withhold the current round, wait for additional sibling information, or roll back and regenerate within horizon $H$ using a fresher visible workspace. Selective per-stream withholding is compatible with the same framework, but is not required by the base formulation above.

In PDT, agreement is the gate that decides whether the system may advance the parallel frontier.

A richer alternative would replace the scalar gate with pairwise or graph-structured compatibility scores between streams. We view that extension as important future work, but the scalar readiness formulation is sufficient to define the core synchronized continuation protocol.

\subsection{Inference-Time Coordination Loop}


The full inference procedure is:

\begin{enumerate}
    \item Encode the prompt with the frozen trunk.
    \item Run the planner to predict latent plan slots $z_{1:S}$.
    \item Project the latent plan into notes space and publish snapshot 0 on the Dynamic Notes Bus.
    \item For synchronization rounds $v = 1, 2, \ldots$ until all streams terminate:
    \item Expose the visible notes window $\mathcal{W}^{(k)}_v$ to each active stream.
    \item Let each active stream emit $\tau$ provisional tokens using its local cache plus SNC conditioning.
    \item Let each active stream emit a provisional latent note $\widehat{\mathbf{n}}^{(k)}_v$.
    \item Compute coverage $\mathbf{c}^{(k)}_v$ and readiness $r^{(k)}_v$ for each active stream.
    \item If $A_v = 1$, commit the current block and schedule the new notes for publication after delay $\Delta$.
    \item Otherwise, stall, selectively withhold, or roll back within horizon $H$, then regenerate with updated visible context.
    \item Merge committed outputs according to planner ownership, section order, and stream-completion state.
\end{enumerate}
The inference loop makes the architectural claim explicit: PDT is a decode $\rightarrow$ summarize $\rightarrow$ agree $\rightarrow$ commit $\rightarrow$ continue protocol.

\subsection{Parameter-Efficient Curriculum}


Training a coordination mechanism on a frozen trunk is unstable if all modules are enabled at once. PDT therefore uses a staged curriculum:

\begin{itemize}
    \item \textbf{Stage 0 (planner pretraining).} Train the planner head, plan embedding pathway, and initial notes projection against latent planner targets and the snapshot-0 plan contract.
    \item \textbf{Stage 1 (stream bootstrap).} Unfreeze stream adapters and SNC cross-attention so streams learn stream-specific conditioning against fixed teacher workspaces.
    \item \textbf{Stage 2 (bus enablement).} Train note-emission modules so streams begin writing learned latent summaries into the versioned bus while the planner seed remains stable.
    \item \textbf{Stage 3 (commit control).} Train coverage and agreement heads so ownership consistency and continuation sufficiency become part of the generation policy.
\end{itemize}

The accompanying dataset and supervision contract reflect the same runtime assumptions. Snapshot 0 is explicitly materialized from the planner contract, canonical plan items are hashed into the shared latent vocabulary, and language-model supervision is masked so each stream is trained only on its assigned section.
