\section{Planner-Seeded Latent Coordination}
\label{sec:snc}

To enable parallel coordination without modifying the pre-trained weights of the base language model, we use a planner-seeded latent workspace together with \textbf{Speculative Note Conditioning (SNC)}. SNC is the read/write mechanism over a shared embedding-space state that lets parallel streams condition on each other's progress.

Formally, let the frozen trunk be parameterized by $\theta_{\text{pre}}$. We introduce a lightweight set of trainable parameters $\phi$, consisting of Stream Adapters, Note Heads, and the SNC Cross-Attention mechanism.

\subsection{Planner-Seeded Note State}
Let $\mathbf{H}_x \in \mathbb{R}^{T \times d}$ denote the prompt hidden states produced by the frozen trunk. The planner predicts logits over $S$ latent plan slots and a shared plan vocabulary of size $V_p$:
\begin{equation}
    \mathbf{\Pi} = \mathrm{PlannerHead}(\mathbf{H}_x) \in \mathbb{R}^{S \times V_p}, \qquad
    z_i = \arg\max_v \mathbf{\Pi}_{i,v}.
\end{equation}
In the reported configuration, $S=16$ and $V_p=65{,}536$. The planner always exposes a fixed-width latent state.

The latent slot IDs are re-embedded and projected into notes space,
\begin{equation}
    \mathbf{e}_i = E_{\text{plan}}[z_i], \qquad
    \mathbf{n}^{\text{plan}}_0 = \mathrm{Norm}\!\left(\frac{1}{|M|} \sum_{i \in M} W_{\text{plan}} \mathbf{e}_i\right),
\end{equation}
where $M$ denotes the active planner slots. The pooled seed $\mathbf{n}^{\text{plan}}_0$ is then broadcast as snapshot 0 on the Dynamic Notes Bus for every stream before decoding begins.
This fixed latent seed couples task decomposition, initial stream ownership, and later coverage supervision before any stream writes free-running content.

\subsection{Frozen Trunk and Stream Adapters}
Let $\mathbf{H}^{(k)}_l \in \mathbb{R}^{T \times d}$ denote the hidden states of stream $k$ at layer $l$. In a standard Transformer, $\mathbf{H}^{(k)}_{l+1} = \text{Block}_{\theta}(\mathbf{H}^{(k)}_l)$. In PDT, we inject stream-specific conditioning via \textbf{Stream Adapters}.

The adapter for stream $k$, denoted as $A_{\phi}^{(k)}$, is a bottleneck MLP with residual connection:
\begin{equation}
    \mathbf{H}^{(k)}_{l, \text{adapt}} = \mathbf{H}^{(k)}_l + \mathbf{W}_{\text{up}} \cdot \sigma(\mathbf{W}_{\text{down}} \cdot \text{LayerNorm}(\mathbf{H}^{(k)}_l))
\end{equation}
where $\sigma$ is the activation function (GELU). These adapters allow the frozen trunk to process identical tokens differently depending on the stream index, breaking the symmetry of parallel decoding and letting multiple streams condition on the same planner seed while specializing into different sections of the output.

\subsection{SNC Cross-Attention Mechanism}
\label{sec:snc_attention}

The core coordination primitive is the Speculative Note Conditioning layer. Unlike standard decoding where context is strictly local, SNC allows stream $k$ to attend to a lagged window of note embeddings assembled from versioned snapshots across sibling streams.

Let $\mathcal{W}^{(k)}_t \in \mathbb{R}^{M_t \times d_{\text{note}}}$ denote the notes window visible to stream $k$ at decode step $t$. The window is built from the most recent lagged snapshots under reveal delay $\Delta$ and a fixed snapshot budget. Reads happen at every decode step; cadence affects when new snapshots are written, not whether the window is consulted. Every generated token is therefore conditioned on a planner-seeded latent workspace even when no new write occurs at that step.

\paragraph{Query-Key-Value Construction.}
The query $\mathbf{Q}^{(k)}$ is derived from the current stream's hidden state. The keys $\mathbf{K}^{(j)}$ and values $\mathbf{V}^{(j)}$ are projected from the sibling notes:
\begin{align}
    \mathbf{Q}^{(k)} &= \mathbf{H}^{(k)}_{l} \mathbf{W}_Q \\
    \mathbf{K}^{(k)}_t &= \mathcal{W}^{(k)}_t \mathbf{W}_K, \quad \mathbf{V}^{(k)}_t = \mathcal{W}^{(k)}_t \mathbf{W}_V
\end{align}
where $\mathbf{W}_{\{Q,K,V\}} \in \phi$ are learned projections.

\paragraph{Trust-Gated Residual Injection.}
A critical challenge in adding attention to a frozen model is preserving the signal magnitude distribution. PDT uses a learned residual gate on the SNC delta before it is added back to the hidden state.

The context vector $\mathbf{C}^{(k)}_t$ is computed via standard scaled dot-product attention over the visible notes window. The final update to the trunk's hidden state is:
\begin{equation}
    \tilde{\mathbf{H}}^{(k)}_{l,t} = \mathbf{H}^{(k)}_{l,t} + \lambda_l \cdot \mathbf{C}^{(k)}_t \mathbf{W}_O
\end{equation}
where $\lambda_l \in [0, 1]$ is the learned residual gate for the instrumented layer. This lets the model retain its pre-trained behavior early in training while gradually incorporating cross-stream information.

\paragraph{Writes and Coverage.}
When cadence logic schedules an update, the speculation head emits a new note embedding
\begin{equation}
    \widehat{\mathbf{n}}^{(k)}_{v+1} = \mathrm{SpeculationHead}\!\left(\tilde{\mathbf{H}}^{(k)}_{l,t}\right),
\end{equation}
which is appended to the versioned bus with stride and version metadata. Coverage is computed separately as
\begin{equation}
    \mathbf{c}^{(k)}_t = \mathrm{CoverageHead}\!\left(\tilde{\mathbf{H}}^{(k)}_{l,t}, E_{\text{plan}}(z), m_{\text{plan}}\right),
\end{equation}
where the plan embeddings come from the same latent vocabulary used by the planner. Because the coverage targets are stream-qualified canonical plan items, they diagnose whether the shared latent plan is being executed coherently and without redundant cross-stream claims.

\subsection{Agreement and Rollback Control}
\label{sec:agreement}

While the SNC attention mechanism enables information flow, it does not guarantee correctness. A sibling stream may still write a misleading summary. To manage this, the \textbf{Agreement Head} is a scalar classifier trained to estimate whether the latest committed token block should be trusted.

For a hidden state $\mathbf{h}_t$, the Agreement Head predicts a trust score $s_t \in [0, 1]$:
\begin{equation}
    s_t = \sigma(\mathbf{w}_{\text{agree}}^T \cdot \text{Dropout}(\mathbf{h}_t) + b_{\text{agree}})
\end{equation}
During inference, this score serves as the rollback trigger. If $s_t < \tau$, the system identifies a coherence failure inside the current commit horizon and regenerates that segment with fresher note context.

This separation of concerns---planner for initial decomposition, SNC for information flow, agreement for control flow, and coverage for plan-item tracking---distributes the coordination policy across the stack rather than concentrating it in any single head.
