\section{Evaluation Scope}
\label{sec:experiments}

The coordination module described in this paper was materially rewritten to match the planner-first latent contract in Sections \ref{sec:architecture} and \ref{sec:snc}. For that reason, quantitative results from earlier module versions are omitted here: they no longer represent the current system. This version of the pre-print is therefore an architecture and implementation paper rather than a settled benchmark paper.

\subsection{What Should Be Evaluated}
The revised system should be evaluated as a coordination mechanism, not merely as a text generator. The relevant question is whether a mandatory prompt-time planner and embeddings-only Dynamic Notes Bus help parallel streams maintain section ownership, avoid redundant content, and remain mutually consistent while decoding.

\subsection{Evaluation Protocol for the Revised Stack}
The current dataset and training pipeline expose the right supervision contract for that question. Each example carries a teacher plan, per-stream \texttt{section\_contract} and \texttt{notes\_contract} fields, a \texttt{plan\_contract} materialized as \texttt{versioned\_notes[0]}, fixed-width \texttt{planner\_ids}, and canonical \texttt{plan\_item\_ids} derived from the same stream-qualified plan catalog. This contract enables evaluation along at least four axes:
\begin{itemize}
    \item planner-slot accuracy against canonical latent planner IDs;
    \item plan-item coverage and ownership consistency against the canonical catalog;
    \item cross-stream overlap and contradiction rates in generated outputs;
    \item latency and throughput relative to serial decoding and external decomposition baselines.
\end{itemize}

\subsection{Current Scope}
This pre-print does not report benchmark numbers for those quantities yet. In particular, it does not claim validated answer-quality gains, latency gains, exact serial equivalence, or superiority over external multi-agent baselines for the revised coordination stack. It also does not claim that the \texttt{planner\_head} alone carries the revised coordination behavior: head-level attribution requires dedicated ablations and intervention studies beyond the current mechanism paper. Quantitative reevaluation on the current implementation is deferred to a future revision.
